Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-09-07 apj.tex
refining things

Non-trivial residual magnitude dispersions are captured in $C_c$.  The square
root of the diagonal elements are denoted as $\sigma$, which
range from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.
Of particular interest within $C_c$ is intrinsic color dispersion.  The intrinsic dispersions in $V$ and colors with respect to $V$ are
\begin{multline}
\sigma(V,U-V,B-V,V-R,V-I)=\\
%\begin{pmatrix}
%\begin{array}{rrrrr}
\left(
0.021^{+0.004}_{-0.005} ,
0.067^{+0.010}_{-0.010},
0.034^{+0.005}_{-0.005},
0.023^{+0.002}_{-0.002},
0.053^{+0.004}_{-0.004}
\right)
%\end{array}
% \end{pmatrix} 
 \text{mag}.
 \label{sig_intrinsic:eqn}
 \end{multline}
The correlations between these parameters are
\begin{multline}
Cor(V,U-V,B-V,V-R,V-I)=\\
\begin{pmatrix}
\begin{array}{rrrrr}
1 & -0.48^{+0.20}_{-0.18} & -0.34^{+0.21}_{-0.22} & 0.88^{+0.10}_{-0.25} & 0.60^{+0.11}_{-0.19} \\
-0.48^{+0.20}_{-0.18} & 1 & 0.60^{+0.10}_{-0.13} & -0.38^{+0.18}_{-0.17} & -0.33^{+0.15}_{-0.14} \\
-0.34^{+0.21}_{-0.22} & 0.60^{+0.10}_{-0.13} & 1 & -0.21^{+0.17}_{-0.17} & 0.08^{+0.14}_{-0.15} \\
0.88^{+0.10}_{-0.25} & -0.38^{+0.18}_{-0.17} & -0.21^{+0.17}_{-0.17} & 1 & 0.65^{+0.06}_{-0.07} \\
0.60^{+0.11}_{-0.19} & -0.33^{+0.15}_{-0.14} & 0.08^{+0.14}_{-0.15} & 0.65^{+0.06}_{-0.07} & 1 \\
\end{array}
 \end{pmatrix}.
 \label{cor_intrinsic:eqn}
 \end{multline}
These color variances are smaller than those derived by \citet{2003A&A...404..901N} or \citet{2007ApJ...659..122J}, for cases where
a direct comparison can be made.
%model11_rubin.py
Having no intrinsic color dispersion is disfavored, and not allowing for them can have important consequences.  Running an analysis assuming no intrinsic dispersion (replacing the probabilistic intrinsic magnitudes in Eqn.~\ref{ewsiv:eqn}
with a strict equality), we find significant shifts in all slope parameters  and smaller statistical uncertainties.

@@

\citet{2017ApJ...842...93M} present an instructive simulation, producing magnitudes and colors of a set of
supernovae 
with an intrinsic source of color variability having relatively smaller value of $\beta$
and a realistic distribution of dust extinction having a relatively larger $\beta$ value.  The expectation is for
supernovae encountering no interstellar dust  to tend to be bluer and exhibit the intrinsic $\beta$ slope for the color-magnitude
relation.  Supernovae encountering thick dust tend to be redder and exhibit the larger dust $\beta$ slope.
The expectation is borne out in their Figure 3, which shows 
observed color versus light-curve shape corrected magnitudes, and is seen in their data Figure 6.
The distinctive expectations manifest in the plot  for the above model allow
the authors to use  information inherent in the distributions to make distinct measurements of intrinsic and dust slopes (referred
therein as $\beta$ using the terminology of SALT2),
despite being
restricted to using one color observable that is  affected by two sources of extinction.

Our Model~II has for each supernova three color parameters that can contain information on more
than one source of color variablility. This allows us to construct from our fit results per-supernova magnitude corrections as a function
of color, to produce a model-based figure analogous to Figure~3 of \citet{2017ApJ...842...93M}.
Such a result is shown in Figure~\ref{speccorr:fig}, which plots
$U$-band magnitude corrections from the three color parameters $\gamma^0_{U} k_0 +\gamma^1_{U} k_1 + \delta_U D$
as a function of $E_o(B-V)$, where the expected color accounts for spectral-feature corrections.
The shallower slope at blue color transitioning to a larger slope at red color 
is correspondence with the expectation of a mixed intrinsic--dust model of the form considered by  \citet{2017ApJ...842...93M}.

@@

If  $\delta D$  is intrinsic, it can be attributed to an intrinsic supernova property  or a dust not described by $\gamma k$.
To first order $\pmb{\delta}$ behaves like standard dust;   $>98$\% of its length projects onto the expectation of  \citetalias{1999PASP..111...63F}
with constant $R^F$.
Therefore, a straightforward explanation is that it describes the absorption due to some form of dust
that does not exactly obey the  \citetalias{1999PASP..111...63F} model.
The departure from   \citetalias{1999PASP..111...63F}
can be seen qualitatively; the decrease in magnitude change toward redder wavelengths, seen in Figure~\ref{deltaratio:fig}, matches the 
expectation for the circumstellar form of dust.
However, we
lack the signal-to-noise to show consistency with this explanation.
The best-fit elements $\delta_X$ do not all share the same sign, as would have been expected for dust.
Confining ourselves to the 98\% of   $\pmb{\delta}$ that lies on the dust plane, associating 
$D$ with $A_V$ would imply an unphysical $R^F=0.5$.
We do not have conclusive evidence that  $\delta$ is being a property of dust distinct from that described by $\gamma$, rather
than representing intrinsic supernova color variation.
\color{black}

>>>>

Non-trivial residual magnitude dispersions are captured in $C_c$.  The square
root of the diagonal elements are denoted as $\sigma$, which
range from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.
Of particular interest within $C_c$ is intrinsic color dispersion.  The intrinsic dispersions in $V$ and colors with respect to $V$ are
\begin{multline}
\sigma(V,U-V,B-V,V-R,V-I)=\\
%\begin{pmatrix}
%\begin{array}{rrrrr}
\left(
0.021^{+0.004}_{-0.005} ,
0.067^{+0.010}_{-0.010},
0.034^{+0.005}_{-0.005},
0.023^{+0.002}_{-0.002},
0.053^{+0.004}_{-0.004}
\right)
%\end{array}
% \end{pmatrix} 
 \text{mag}.
 \label{sig_intrinsic:eqn}
 \end{multline}
The correlations between these parameters are
\begin{multline}
Cor(V,U-V,B-V,V-R,V-I)=\\
\begin{pmatrix}
\begin{array}{rrrrr}
1 & -0.48^{+0.20}_{-0.18} & -0.34^{+0.21}_{-0.22} & 0.88^{+0.10}_{-0.25} & 0.60^{+0.11}_{-0.19} \\
-0.48^{+0.20}_{-0.18} & 1 & 0.60^{+0.10}_{-0.13} & -0.38^{+0.18}_{-0.17} & -0.33^{+0.15}_{-0.14} \\
-0.34^{+0.21}_{-0.22} & 0.60^{+0.10}_{-0.13} & 1 & -0.21^{+0.17}_{-0.17} & 0.08^{+0.14}_{-0.15} \\
0.88^{+0.10}_{-0.25} & -0.38^{+0.18}_{-0.17} & -0.21^{+0.17}_{-0.17} & 1 & 0.65^{+0.06}_{-0.07} \\
0.60^{+0.11}_{-0.19} & -0.33^{+0.15}_{-0.14} & 0.08^{+0.14}_{-0.15} & 0.65^{+0.06}_{-0.07} & 1 \\
\end{array}
 \end{pmatrix}.
 \label{cor_intrinsic:eqn}
 \end{multline}
These color variances are smaller than those derived by \citet{2003A&A...404..901N} or \citet{2007ApJ...659..122J}, for cases where
a direct comparison can be made.
\color{red}
\citep{2017ApJ...842...93M}  find an intrinsic dispersion of the SALT2 $c$ parameter (associated with $B-V$) of $\approx 0.07$ mag.
\color{black}
%model11_rubin.py
Having no intrinsic color dispersion is disfavored, and not allowing for them can have important consequences.  Running an analysis assuming no intrinsic dispersion (replacing the probabilistic intrinsic magnitudes in Eqn.~\ref{ewsiv:eqn}
with a strict equality), we find significant shifts in all slope parameters  and smaller statistical uncertainties.

@@

\citet{2017ApJ...842...93M} discuss an instructive simulation
that considers a set of
supernovae 
having an intrinsic linear color-magnitude relationship with slope $\beta$.
(The absence of an intrinsic color-magnitude relation corresponds to $\beta=0$.)
Each supernova encounters dust with a relatively larger slope
for the color-magnitude relationship and
extinction drawn from a realistic extinction distribution.
The expectation is for
supernovae encountering no interstellar dust  to tend to be bluer and
collectively exhibit the intrinsic $\beta$ slope for the color-magnitude
relation.  Supernovae encountering thick dust tend to be redder and collectively exhibit the larger dust  slope.
This expectation is borne out in Figure 3 of \citet{2017ApJ...842...93M}, which
shows  simulated
observed color versus light-curve shape corrected magnitudes.
The distinctive distribution of supernovae in color-magnitude space  allows
the authors to distinguish intrinsic $\beta$ and the dust slope,
despite being
restricted to using one color observable that is  affected by two sources of extinction.

In our Model~II, each supernova has three color parameters that can contain information on more
than one source of color variablility. This allows us to construct from our fit results per-supernova magnitude corrections as a function
of color, to produce a model-based figure analogous to Figure~3 of \citet{2017ApJ...842...93M}.
Such a result is shown in Figure~\ref{speccorr:fig}, which plots
$U$-band magnitude corrections from the three color parameters $\gamma^0_{U} k_0 +\gamma^1_{U} k_1 + \delta_U D$
as a function of $E_o(B-V)$, where the expected color accounts for spectral-feature corrections.
The shallower (but non-zero) slope at blue color transitioning to a larger slope at red color 
is correspondence with the expectation of a mixed intrinsic--dust model.
The fact that
, and is seen in their data Figure 6 of the form considered by  \citet{2017ApJ...842...93M}.
Slope is not zero.

@@

Considering  $\delta D$  to be intrinsic, it can be attributed to an intrinsic supernova property  or a dust not described by $\gamma k$.
To first order $\pmb{\delta}$ behaves like standard dust;   $>98$\% of its length projects onto the expectation of  \citetalias{1999PASP..111...63F}
with constant $R^F$.
Therefore, a straightforward explanation is that it describes the absorption due to some form of dust
that does not exactly obey the  \citetalias{1999PASP..111...63F} model.
The departure from   \citetalias{1999PASP..111...63F}
can be seen qualitatively; the decrease in magnitude change toward redder wavelengths, seen in Figure~\ref{deltaratio:fig}, matches the 
expectation for the circumstellar form of dust.
However, we
lack the signal-to-noise to show consistency with this explanation.
The best-fit elements $\delta_X$ do not all share the same sign, as would have been expected for dust.
Confining ourselves to the 98\% of   $\pmb{\delta}$ that lies on the dust plane, associating 
$D$ with $A_V$ would imply an unphysical $R^F=0.5$.
We do not have conclusive evidence that  $\delta$ is being a property of dust distinct from that described by $\gamma$, rather
than representing intrinsic supernova color variation.
\color{black}
